---
title: "Run 'Ri.NAmerica' model trained on SLF invaded region in North America"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu, [ORCID](https://orcid.org/0009-0001-2338-7928)]"
date: "2026-07-30"
---

# Overview

In the last vignette, I cropped the covariate layer to the extent of each regional model training area and chose background points from those layers to characterize these covariates. In this vignette, I will train our first regional-scale model based on the invaded North American range. This another step toward my goal of analyzing the realized niche of Lycorma delicatula at multiple spatial scales. This output will be combined with the other regional models and compared directly to the global scale model to determine the impact of climate change on the suitability of SLF in important viticultural regions.

Just to reiterate, I hypothesize that by creating models at multiple spatial scales, I can have more confidence in suitability predictions made for specific SLF populations and important viticultural regions, and can predict the risk of specific populations spreading now and under climate change. I also hypothesize that the regional-scale models, once ensembled into a single prediction, will make more refined predictions than a global-scale model could.

See vignette 050 where I ran the global model for details on the reasoning behind my choices in analyses within this vignette.

# Setup

*CURRENT MODEL VERSION: v9*

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
# here() starts at the root folder of this package.
library(devtools)
library(knitr)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# spatial data handling
library(raster) 
library(terra) 

library(viridis)

library(scari)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

`SDMtune` will run MaxEnt through java via the `rJava` package. You will need to ensure that your machine has the proper version of java installed (x32 or x64) for your operating system.

```{r check maxent installation, message = FALSE}

checkMaxentInstallation(verbose = TRUE)

```

This chunk sets the java memory allocation (`Xmx`). I will increase the memory allocation from 512m (the default) to 8GB (8000mb) of memory. This should be edited according to your system, because you should still leave about 20% of your system memory available for other processes. If you have a 32-bit system, you cannot set the max above 4000mb.

```{r control MaxEnt settings}

# xmx sets java memory allocation
options(java.parameters = "-Xmx8000m")

# xss sets java stack size
# options(java.parameters = c("-Xss2560k", "-Xmx2048m"))

```

# 1. Format Data for Model

I will load in the datasets I will need for the MaxEnt models. These are labeled at the beginning of the object name by the parameter they will be used in the `SDMtune::train()` function (x, p or a). I will begin by loading in the covariate data and then by loading in the points datasets. The `env_covariates` dataset labeled `invaded_n_american` contains a raster of each of the environmental covariates used to train the model- these were the covariates that were masked using the K-G climate zones in the last vignette.

## 1.1 Input Data- env covariates

```{r load in historical covariates}

# path to directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")


# the env covariate scale used to train the model
x_invaded_n_american_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent_10km"), pattern = "\\_regional_invaded_n_american_KG.asc$", full.names = TRUE)  %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# the scale used to make xy predictions
x_global_hist_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent_10km"), pattern = "\\.asc$", full.names = TRUE) %>%
    grep("bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

```

The CMIP6 versions of these covariates will only be used for projection purposes.

```{r load in CMIP6 covariates}

# path to directory
  mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/future_climate_rasters/chelsa2.1_30arcsec")


# the env covariates for performing xy predictions for global slf and IVR points

# SSP126
x_global_126_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp126_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# SSP370
x_global_370_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp370_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

# SSP585
x_global_585_env_covariates_list <- list.files(path = file.path(mypath, "2041-2070_ssp585_GFDL", "v1_maxent_10km"), pattern = "\\_global.asc$", full.names = TRUE) %>%
  # dont include Access to cities
  grep(pattern = "atc_2015", value = TRUE, invert = TRUE)

```

I will create rasters of the environmental covariates, stack and gather summary statistics. I will also shorten their names and exclude possible operators from layer names (for example, using the dash symbol was found to interfere with SDMtune making predictions for tables downstream).

```{r naming object}

# layer name object. Check order of layers first
env_layer_names <- c("bio11", "bio12", "bio15", "bio2")

```

```{r stack historical covariates and make naming consistent, message = FALSE}

# stack env covariates
x_invaded_n_american_env_covariates <- terra::rast(x = x_invaded_n_american_env_covariates_list)

# attributes
nlyr(x_invaded_n_american_env_covariates)
names(x_invaded_n_american_env_covariates)
minmax(x_invaded_n_american_env_covariates)
# ext(x_invaded_n_american_env_covariates)
# crs(x_invaded_n_american_env_covariates)

# I will change the name of the variables because they are throwing errors in SDMtune
names(x_invaded_n_american_env_covariates) <- env_layer_names



# global rasters
# stack env covariates
x_global_hist_env_covariates <- terra::rast(x = x_global_hist_env_covariates_list)

# attributes
nlyr(x_global_hist_env_covariates)
names(x_global_hist_env_covariates)
minmax(x_global_hist_env_covariates)
# ext(x_global_hist_env_covariates)
# crs(x_global_hist_env_covariates)

# I will change the name of the variables because they are throwing errors in SDMtune
names(x_global_hist_env_covariates) <- env_layer_names
# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)


```

```{r stack CMIP6 covariates, message = FALSE}

# SSP126
x_global_126_env_covariates <- terra::rast(x = x_global_126_env_covariates_list)

# attributes
nlyr(x_global_126_env_covariates)
names(x_global_126_env_covariates)
minmax(x_global_126_env_covariates)
# ext(x_global_126_env_covariates)
# crs(x_global_126_env_covariates)

names(x_global_126_env_covariates) <- env_layer_names


# SSP370
x_global_370_env_covariates <- terra::rast(x = x_global_370_env_covariates_list)

# attributes
nlyr(x_global_370_env_covariates)
names(x_global_370_env_covariates)
minmax(x_global_370_env_covariates)
# ext(x_global_370_env_covariates)
# crs(x_global_370_env_covariates)

names(x_global_370_env_covariates) <- env_layer_names


# SSP585
x_global_585_env_covariates <- terra::rast(x = x_global_585_env_covariates_list)

# attributes
nlyr(x_global_585_env_covariates)
names(x_global_585_env_covariates)
minmax(x_global_585_env_covariates)
# ext(x_global_585_env_covariates)
# crs(x_global_585_env_covariates)

names(x_global_585_env_covariates) <- env_layer_names

```

```{r remove lists}

rm(x_invaded_n_american_env_covariates_list)
rm(x_global_hist_env_covariates_list)
rm(x_global_126_env_covariates_list)
rm(x_global_370_env_covariates_list)
rm(x_global_585_env_covariates_list)

```

## 1.2 Input data- presences / absences (training and testing)

I need to also load in the SLF presence and background points datasets created in vignette 060.

```{r load in points datasets}

# slf presences
p_slf_points <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2026-07-30.csv")) %>%
  dplyr::select(-species)

# background points
a_regional_invaded_n_american_background_points <- read.csv(file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_background_points_v9.csv"))

```

I will need to combine the above covariates, presence and background point datasets into two different datasets to feed into MaxEnt: one for training and another for testing the model. These dataset will need to contain point-wise values for each of the predictor covariates at the presence and background point coordinates. `SDMtune` takes an SWD (sample with data) S4 object for these purposes, containing the presences and background points with associated covariate data to be fed into the model. First, I will create the training dataset. I will need to perform some extra edits on the slf presences. These presences need to be filtered into the training and testing datasets. I will filter the presences from the invaded range in North America and will use these points to train the invaded model, while the presences from the invaded range in Asia will be used to validate / test the model.

### Model training object

```{r import training points}

p_slf_invaded_n_american_train <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_train_slf_presences_v9.csv"))

```

Now I will create the "samples with data" object that SDMtune requires.

```{r prepare training SWD data object}

regional_invaded_n_american_train <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_invaded_n_american_env_covariates,
  p = p_slf_invaded_n_american_train, 
  a = a_regional_invaded_n_american_background_points,
  verbose = TRUE # print helpful messages
  )

regional_invaded_n_american_train@coords # coordinates
regional_invaded_n_american_train@pa # presence / absence (background counted as absence)
regional_invaded_n_american_train@data # extracted data from 


# OLD method
#regional_trainTest <-  SDMtune::trainValTest(
#  x = regional_invaded_n_american_SWD,
#  test = 0.2,
#  only_presence = TRUE,
#  seed = 6
#)

```

I usually look at how many records were dropped before I save it, because SDMtune gives an undefined warning about how many samples were discarded. In this case, we only lost 3 presence records, so this is acceptable. I will also save the output to be used later.

```{r save training SWD file}

SDMtune::swd2csv(swd = regional_invaded_n_american_train, file_name = c(
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_train_slf_presences_with_data_v9.csv"),
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_background_points_with_data_v9.csv")
  ))

```

### Model testing object

The testing / validation SWD object will be created in the same fashion, except using a subset of the presences from the invaded range in Asia. These presences will be selected using a shapefile of the Japan/SK extent as a mask. we chose to test using the invaded region over the native region because Sobek-Swant et al found that their invaded MaxEnt models performed poorly when projected to the native range for EAB. Additionally, the initial versions of this model performed very poorly when tested on the native range (Sobek-Swant et al, 2012).

```{r import test points}

p_slf_invaded_n_american_test <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_test_slf_presences_v9.csv"))

```

```{r prepare testing SWD data object}

regional_invaded_n_american_test <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_global_hist_env_covariates, 
  p = p_slf_invaded_n_american_test, 
  a = a_regional_invaded_n_american_background_points,
  verbose = TRUE # print helpful messages
  )

regional_invaded_n_american_test@coords # coordinates
regional_invaded_n_american_test@pa # presence / absence (background counted as absence)
regional_invaded_n_american_test@data # extracted data from 

```

```{r save testing SWD file}

SDMtune::swd2csv(swd = regional_invaded_n_american_test, file_name = c(
  file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_test_slf_presences_with_data_v9.csv"),
  file.path(here::here(), "vignette-outputs", "data-tables", "NULL.csv")
  ))

# remove second copy of background data
file.remove(file.path(here::here(), "vignette-outputs", "data-tables", "NULL.csv"))

```

# 2. Train Invaded Regional Model

First, I will train a MaxEnt model. This and the other regional-scale models will NOT be cross-validated. Cross-validation randomly selects a percentage of the presence data for training and testing, without regard for spatial location. This process makes sense for the global model, which is fed ALL presence data. Due to the spatial scale of the regional models, it makes more sense to spatially separate the training and testing data if given the opportunity (plus this is considered "best practice")

I will use the following list of hyperparameters to train the initial model, which I selected via tuning for the global model.

-   ALL feature classes (fc) used (l = linear, q = quadratic, p = product, h = hinge, t = threshold)
-   regularization multiplier (reg) set to 1.5 (more regularized than default of 1)
-   iterations = 5000. This is the max number of iterations for the optimization algorithm to perform before stopping training. Increasing this number from the default of 500 allows the algorithm to make more refined predictions.

```{r train maxent model}

regional_invaded_n_american_model <- SDMtune::train(
  method = "Maxent",
  data = regional_invaded_n_american_train,
  fc = "qpht", # feature classes set to ALL
  reg = 1.5,
  iter = 5000, # number of iterations
  progress = TRUE
)

```

## Summary Statistics

This function produces all summary statistics for this model. For the complete and annotated workflow used to create this function, see `051_compute_MaxEnt_summary_statistics_workflow.Rmd`.

```{r compute summary statistics using custom function}

# output directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models/slf_regional_invaded_n_american_v9")

scari::compute_MaxEnt_summary_statistics(
  model.obj = regional_invaded_n_american_model, 
  model.name = "regional_invaded_n_american", 
  mypath = mypath, 
  create.dir = TRUE, # create subdirectory
  env.covar.obj = x_invaded_n_american_env_covariates, # env covariates raster stacked
  train.obj = regional_invaded_n_american_train, # training data used to create model
  test.obj = regional_invaded_n_american_test, # data you wish to use to test the model
  plot.type = "cloglog", # types of univariate and marginal response curves to be created
  jk.test.type = c("train", "test") # types of jackknife curves to be created
  )

```

Looking at the summary statistics, the most explanatory variable (according to the jackknife test, training data in North America) is the seasonal precipitation seasonality (Bio 15). By itself, it can produce nearly the same AUC as the full model. However, when left out, Bio 11 (minimum winter temperature) has the most negative impact on the overall model (so it is the most explanatory). Meanwhile, for the testing data in the Asian invaded range, Bio 11 is overall the most impactful variable. It produces nearly the same model by itself, while affecting the model most heavily when left out. This aligns with the literature, showing that SLF are heavily influenced by winter temperature minimums in Korea (Lee et al, 2011). The permutation importance plot validates these points.

Here are the jackknife plots:
```{r load jackknife plots, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# training
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_jackknife_training.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

# training
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_jackknife_testing.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

The univariate response curves (the result of creating a model with only that variable) resemble those from the the global scale mean model. Bio 11 resembles a bell-shaped curve, with a peak around 0°C. The activity of Bio 15 suggests a threshold that trails off from its max at about 10% precipitation seasonality. Bio 12 is a sharp peak at around 1500mm precipitation, and Bio 2 generally has activity between 4 and 10.5°C.

Here are the response curves:
```{r load response curve plots, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# bio 2
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_bio2_univar_resp_curve_cloglog.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

# bio 11
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_bio11_univar_resp_curve_cloglog.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

# bio 12
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_bio12_univar_resp_curve_cloglog.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

# bio 15
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_bio15_univar_resp_curve_cloglog.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

The model itself had a training AUC of 0.977 and a test AUC of 0.543. The test AUC is less than desired, but it is higher than 0.5 (the threshold at which the model does not discern better than random). It is likely overfit to the training points because of the high number of training points within a confined geographical area. This is due to the higher sampling effort for SLF in the eastern USA than anywhere else in the world.  However, it is highly likely that SLF has not yet reached all suitable areas within North America, given the yearly trend of range expansion, so this model will likely not predict as well as it would if the species were at equilibrium.

The overfit nature of this model is not an issue; our pipeline is designed to average the outputs of multiple models together for more robust predictions. Each model will be weighted by its AUC, so this model will not contribute as much to the ensemble of regional models as one with a higher fit.  

Here is the ROC curve:
```{r load ROC plots, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# ROC
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_ROC.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

# 3. Create Outputs for Analysis

## 3.1 Create distribution map for area of interest

Lastly, I will use the `SDMtune::predict()` function to predict the suitability for the range of interest. I will threshold the map by the `fixed_1`, `MTSS` and `10_percentile` thresholds.

```{r set wd}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models/slf_regional_invaded_n_american_v9")

```

```{r re-load model as needed}
  
regional_invaded_n_american_model <- read_rds(file = file.path(mypath, "regional_invaded_n_american_model.rds"))

```

```{r predict mean and thresholded suitability maps for historical data}

scari::create_MaxEnt_suitability_maps(
  model.obj = regional_invaded_n_american_model,
  model.name = "regional_invaded_n_american", 
  mypath = mypath, 
  create.dir = FALSE, 
  env.covar.obj = x_global_hist_env_covariates, 
  describe.proj = "globe_1981-2010", # name of area or time period describe.proj to
  clamp.pred = TRUE,
  thresh = c("fixed_1", "10_percentile", "MTSS"),
  map.thresh = TRUE, # whether thresholded versions of these maps should be created
  map.thresh.extra = "fixed_1",
  summary.file = file.path(mypath, "regional_invaded_n_american_summary.csv")
)

```

I will also create suitability maps for ssp scenarios of the projected 2041-2070 climate data.

<!--

```{r predict suitability maps for ssp126 data}

scari::create_MaxEnt_suitability_maps(
  model.obj = regional_invaded_n_american_model,
  model.name = "regional_invaded_n_american", 
  mypath = mypath, 
  create.dir = FALSE, 
  env.covar.obj = x_global_126_env_covariates, 
  describe.proj = "globe_2041-2070_GFDL_ssp126", # name of area or time period projected to
  clamp.pred = TRUE,
  thresh = c("fixed_1", "10_percentile", "MTSS"),
  map.thresh = TRUE, # whether thresholded versions of these maps should be created
  map.thresh.extra = "fixed_1",
  summary.file = file.path(mypath, "regional_invaded_n_american_summary.csv")
)

```

```{r predict suitability maps for ssp370 data}

scari::create_MaxEnt_suitability_maps(
  model.obj = regional_invaded_n_american_model,
  model.name = "regional_invaded_n_american", 
  mypath = mypath, 
  create.dir = FALSE, 
  env.covar.obj = x_global_370_env_covariates, 
  describe.proj = "globe_2041-2070_GFDL_ssp370", # name of area or time period projected to
  clamp.pred = TRUE,
  thresh = c("fixed_1", "10_percentile", "MTSS"),
  map.thresh = TRUE, # whether thresholded versions of these maps should be created
  map.thresh.extra = "fixed_1",
  summary.file = file.path(mypath, "regional_invaded_n_american_summary.csv")
)

```

```{r predict suitability maps for ssp585 data}

scari::create_MaxEnt_suitability_maps(
  model.obj = regional_invaded_n_american_model,
  model.name = "regional_invaded_n_american", 
  mypath = mypath, 
  create.dir = FALSE, 
  env.covar.obj = x_global_585_env_covariates, 
  describe.proj = "globe_2041-2070_GFDL_ssp585", # name of area or time period projected to
  clamp.pred = TRUE,
  thresh = c("fixed_1", "10_percentile", "MTSS"),
  map.thresh = TRUE, # whether thresholded versions of these maps should be created
  map.thresh.extra = "fixed_1",
  summary.file = file.path(mypath, "regional_invaded_n_american_summary.csv")
)

```

-->

The suitability maps for this model are shown below. I will load in the maps that were thresholded by both the MTSS and MTP thresholds. Bear in mind that I chose the MTSS threshold for analyses downstream.

The most prominent suitable areas above the MTSS (the colored areas) were predicted in North America and Europe. Notably, this model omits known suitable area in China, Japan and Korea, so it is not the most accurate in it those predictions. It does resemble the global-scale model in its predictions for Europe and North America, however smaller. There are also notable suitable areas in New Zealand and southern Argentina within the southern hemisphere. Here are the predicted suitability maps. Notably, as the ssp scenario becomes more severe (ssp126 -> ssp585), the predicted suitable area of each map slowly shifts northwards and becomes smaller. When interpreting this model, bear in mind that we will include it within an ensemble average of models and that its contributions toward that ensemble will be weighted by its AUC (considerably low). However, we will not omit it because it likely contains important trends in climatic suitability (see the summary statistics and the response curve for Bio 15). Here are the suitability maps:

The historical data (1981-2010):
```{r load suitability maps hist, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# load in maps
# historical data
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_pred_suit_clamped_cloglog_thresholded_MTSS_fixed_1_globe_1981-2010.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

The predicted future data (2041-2070) under the ssp126 climate change scenario:
```{r load suitability maps ssp126, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# load in maps
# ssp126 for 2041-2070 data
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_pred_suit_clamped_cloglog_thresholded_MTSS_fixed_1_globe_2041-2070_GFDL_ssp126.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

The predicted future data (2041-2070) under the ssp370 climate change scenario:
```{r load suitability maps ssp370, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# load in maps
# ssp126 for 2041-2070 data
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_pred_suit_clamped_cloglog_thresholded_MTSS_fixed_1_globe_2041-2070_GFDL_ssp370.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

The predicted future data (2041-2070) under the ssp585 climate change scenario:
```{r load suitability maps hssp585ist, eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE}

# load in maps
# ssp126 for 2041-2070 data
knitr::include_graphics(
  path = file.path(mypath, "plots", "regional_invaded_n_american_pred_suit_clamped_cloglog_thresholded_MTSS_fixed_1_globe_2041-2070_GFDL_ssp585.jpg"),
  rel_path = FALSE, # already done
  error = TRUE
  )

```

## 3.2 Predict suitability for known SLF presences

```{r data import- slf populations}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models/slf_regional_invaded_n_american_v9")

# slf presence data
slf_presences <- read.csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "slf_all_coords_final_2026-07-30.csv")) %>%
  dplyr::select(-species)

```

```{r predictions for known SLF populations- historical}

scari::predict_xy_suitability(
  xy.obj = slf_presences,
  xy.type = "Lycorma delicatula",
  env.covar.obj = x_global_hist_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_slf_all_coords_1981-2010"
)

```

I will repeat the same process for the CMIP6 data.

<!--

```{r predictions for known SLF populations- ssp126}

scari::predict_xy_suitability(
  xy.obj = slf_presences,
  xy.type = "Lycorma delicatula",
  env.covar.obj = x_global_126_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_slf_all_coords_2041-2070_GFDL_ssp126"
)

```

```{r predictions for known SLF populations- ssp370}

scari::predict_xy_suitability(
  xy.obj = slf_presences,
  xy.type = "Lycorma delicatula",
  env.covar.obj = x_global_370_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_slf_all_coords_2041-2070_GFDL_ssp370"
)

```

```{r predictions for known SLF populations- ssp585}

scari::predict_xy_suitability(
  xy.obj = slf_presences,
  xy.type = "Lycorma delicatula",
  env.covar.obj = x_global_585_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_slf_all_coords_2041-2070_GFDL_ssp585"
)

```

-->

## 3.3 Predict suitability for IVR regions

I will perform the same action as above, but for the locations of important wineries around the world.

```{r data import- wineries}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models/slf_regional_invaded_n_american_v9")

# load in all IVR points
IVR_regions <- readr::read_rds(file.path(here::here(), "data", "wineries_esri54017.rds"))

IVR_regions <- IVR_regions %>%
  dplyr::select(x, y)

```

```{r predictions for IVR coordinate data- historical}

scari::predict_xy_suitability(
  xy.obj = IVR_regions,
  xy.type = "IVR regions",
  env.covar.obj = x_global_hist_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_wineries_1981-2010",
  buffer.pred = TRUE
)

```

I will do the same for the SSP scenarios.

<!--

```{r predictions for IVR coordinate data- ssp126}

scari::predict_xy_suitability(
  xy.obj = IVR_regions,
  xy.type = "IVR regions",
  env.covar.obj = x_global_126_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_wineries_2041-2070_GFDL_ssp126",
  buffer.pred = TRUE
)

```

```{r predictions for IVR coordinate data- ssp370}

scari::predict_xy_suitability(
  xy.obj = IVR_regions,
  xy.type = "IVR regions",
  env.covar.obj = x_global_370_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_wineries_2041-2070_GFDL_ssp370",
  buffer.pred = TRUE
)

```

```{r predictions for IVR coordinate data- ssp585}

scari::predict_xy_suitability(
  xy.obj = IVR_regions,
  xy.type = "IVR regions",
  env.covar.obj = x_global_585_env_covariates,
  model.obj = regional_invaded_n_american_model,
  mypath = mypath,
  predict.type = "cloglog",
  output.name = "regional_invaded_n_american_wineries_2041-2070_GFDL_ssp585",
  buffer.pred = TRUE
)

```

-->

We have trained the Ri.NAmerica model, so now we will move on to train the other two regional-scale models. 

# References

1. Damus, M. (2014, May 23). Threshold rule [Online post]. Google Groups: Maxent. https://groups.google.com/g/maxent/c/MUMu4KP7kSs?pli=1

2. Elith, J., Phillips, S. J., Hastie, T., Dudík, M., Chee, Y. E., & Yates, C. J. (2011). A statistical explanation of MaxEnt for ecologists: Statistical explanation of MaxEnt. Diversity and Distributions, 17(1), 43–57. https://doi.org/10.1111/j.1472-4642.2010.00725.x

3. Feng, X. (2022). Shandongfx/nimbios_enm [HTML]. https://github.com/shandongfx/nimbios_enm (Original work published 2018).

4. Gallien, L., Douzet, R., Pratte, S., Zimmermann, N. E., & Thuiller, W. (2012). Invasive species distribution models – how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography, 21(11), 1126–1136. https://doi.org/10.1111/j.1466-8238.2012.00768.x

5. Liu, C., Newell, G., & White, M. (2016). On the selection of thresholds for predicting species occurrence with presence-only data. Ecology and Evolution, 6(1), 337–348. https://doi.org/10.1002/ece3.1878

6. Liu, C., White, M., & Newell, G. (2013). Selecting thresholds for the prediction of species occurrence with presence-only data. Journal of Biogeography, 40(4), 778–789. https://doi.org/10.1111/jbi.12058

7. Maxent model Summary Output. (2007, April 11). Maxent Model for Swainson’s Hawk (SWHA). https://geo.pointblue.org/lipmap/maxent_hu_ecoregion/SWHA.html

8. Radosavljevic, A., & Anderson, R. P. (2014). Making better Maxent models of species distributions: Complexity, overfitting and evaluation. Journal of Biogeography, 41(4), 629–643. https://doi.org/10.1111/jbi.12227

9. Phillips, S. J., Anderson, R. P., & Schapire, R. E. (2006). Maximum entropy modeling of species geographic distributions. Ecological Modelling, 190(3), 231–259. https://doi.org/10.1016/j.ecolmodel.2005.03.026

10. Steven Phillips. (2017). A Brief Tutorial on Maxent. http://biodiversityinformatics.amnh.org/open_source/maxent/.

11. Steven J. Phillips, Miroslav Dudík, & Robert E. Schapire. (2023). Maxent software for modeling species niches and distributions (Version 3.4.3 (Java)) [Computer software]. http://biodiversityinformatics.amnh.org/open_source/maxent/.

12. Srivastava, V., Roe, A. D., Keena, M. A., Hamelin, R. C., & Griess, V. C. (2021). Oh the places they’ll go: Improving species distribution modelling for invasive forest pests in an uncertain world. Biological Invasions, 23(1), 297–349. https://doi.org/10.1007/s10530-020-02372-9

13. Valavi, R., Elith, J., Lahoz-Monfort, J. J., & Guillera-Arroita, G. (2019). blockCV: An r package for generating spatially or environmentally separated folds for k-fold cross-validation of species distribution models. Methods in Ecology and Evolution, 10(2), 225–232. https://doi.org/10.1111/2041-210X.13107

14. VanDerWal, J., Shoo, L. P., Graham, C., & Williams, S. E. (2009). Selecting pseudo-absence data for presence-only distribution modeling: How far should you stray from what you know? Ecological Modelling, 220(4), 589–594. https://doi.org/10.1016/j.ecolmodel.2008.11.010

15. Vignali, S., Barras, A. G., Arlettaz, R., & Braunisch, V. (2020). SDMtune: An R package to tune and evaluate species distribution models. Ecology and Evolution, 10(20), 11488–11506. https://doi.org/10.1002/ece3.6786
