---
title: "Assess model goodness of fit via AUC, TSS and confusion matrices"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu, [ORCID](https://orcid.org/0009-0001-2338-7928)]"
date: "2026-08-10"
---

In this vignette, I will assess the fit of my models using a few different metrics: the TSS score, AUC score, and confusion matrices.

First, I will look at the TSS score, which is a threshold-dependent metric of model fit. Considering our decision to convert our risk maps to presence/absence maps using the MTSS threshold, this is an important step. The TSS measures the fit at the threshold where the TSS is maximized (the MTSS threshold).

Next, we will look at the AUC score, which is a threshold-independent metric of model fit. The AUC score is the area under the ROC curve, which is a plot of the true positive rate (sensitivity) against the false positive rate (1 - specificity). The AUC score ranges from 0 to 1, with a score of below 0.5 indicating no discrimination, a score of 0.5 being discrimination equal to a random result, and a score of 1 indicating perfect discrimination. Overall, this test assesses the model's ability to discriminate a true positive from a false positive.I will plot all four ROC curves on the same plot, allowing for a direct comparison of the AUC (goodness of fit) between models.

Finally, I will look at confusion matrices, which provide a summary of the model's performance in terms of true positives, false positives, true negatives, and false negatives. These matrices will allow us to calculate sensitivity, specificity, commission error, and omission error for each model.

# Setup

```{r load necesssary packages, message = FALSE}

# general tools
library(tidyverse)  #data manipulation
library(here) #making directory pathways easier on different instances
# here::here() starts at the root folder of this package.
library(devtools)

# spatial data
library(terra)

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs

# html tools
library(kableExtra)
library(webshot)
library(webshot2)

```

These plots depict the activity of the variables put into the MaxEnt models. I will create a combined figure for the 3 models in the regional_ensemble

```{r set wd 1}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

```{r style object}

ensemble_colors <- c(
  "Rn (native)" = "#4daf4a",
  "Ri.NAmerica" =  "#e41a1c",
  "Ri.Asia" = "#377eb8"
)

```

```{r load model objects}

regional_native_model <- read_rds(file = file.path(mypath, "slf_regional_native_v5", "regional_native_model.rds"))

regional_invaded_n_american_model <- read_rds(file = file.path(mypath, "slf_regional_invaded_n_american_v9", "regional_invaded_n_american_model.rds"))

regional_invaded_asian_model <- read_rds(file = file.path(mypath, "slf_regional_invaded_asian_v4", "regional_invaded_asian_model.rds"))

```

# TSS Score

First, we will assess the TSS score for each model. The TSS score is a measure of the model's ability to discriminate between presence and absence data, with values ranging from -1 to 1. A value of 0 indicates no discrimination, while a value of 1 indicates perfect discrimination. It is a threshold-dependent measure of accuracy, and we have already calculated the TSS score for each model based on the testing data. I will load those in and display them.

```{r import TSS scores, message = FALSE}

regional_native_tss <- read_csv(file = file.path(mypath, "slf_regional_native_v5", "regional_native_model_eval_metrics.csv")) %>%
  # select test TSS
  dplyr::select(TSS) %>%
  dplyr::slice(2)

regional_invaded_n_american_tss <- read_csv(file = file.path(mypath, "slf_regional_invaded_n_american_v9", "regional_invaded_n_american_model_eval_metrics.csv")) %>%
  dplyr::select(TSS) %>%
  dplyr::slice(2)

regional_invaded_asian_tss <- read_csv(file = file.path(mypath, "slf_regional_invaded_asian_v4", "regional_invaded_asian_model_eval_metrics.csv")) %>%
  dplyr::select(TSS) %>%
  dplyr::slice(2)

```

```{r display TSS scores}

tss_scores <- tibble(
  model = c("regional native", "regional invaded North American", "regional invaded Asian"),
  TSS = c(
    regional_native_tss$TSS,
    regional_invaded_n_american_tss$TSS,
    regional_invaded_asian_tss$TSS
  )
)

tss_scores

```

We can see that the native and invaded Asian models have excellent discriminating power, with TSS scores of 0.99 and 0.58, respectively. The invaded North American model has a TSS score of 0.32, indicating that it's discriminating power is poor. 

Despite the poor performance of some of our regional-scale models, we can have confidence in our ensembled mean model, which is composed of the three models put together using a weighted mean of each of the regional-scale models. This gave higher weight to models that performed better according to the AUC metric, and lower weight to the lower performers. Indeed, the native model contributed 46% of the final model output, the invaded Asian model contributed 31% of the final output, and the invaded North American model only contributed 24% of the final weight. This correlates with our TSS scores.

# ROC regional ensemble

## Prepare data

First, I will build the SWD objects required by `SDMtune::plotROC`

```{r load testing points datasets, message = FALSE}

regional_native_test <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_test_slf_presences_v5.csv"))

regional_invaded_n_american_test <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_test_slf_presences_v9.csv"))

regional_invaded_asian_test <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_test_slf_presences_v4.csv"))

```

```{r load in bg points datasets, message = FALSE}

# background points
a_regional_invaded_n_american_background_points <- read.csv(file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_n_american_background_points_v9.csv"))

a_regional_native_background_points <- read.csv(file.path(here::here(), "vignette-outputs", "data-tables", "regional_native_background_points_v5.csv"))

a_regional_invaded_asian_background_points <- read.csv(file.path(here::here(), "vignette-outputs", "data-tables", "regional_invaded_asian_background_points_v4.csv"))

```

```{r load in historical covariates}

# path to directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/historical_climate_rasters/chelsa2.1_30arcsec")


# the scale used to make xy predictions
x_global_hist_env_covariates_list <- list.files(path = file.path(mypath, "v1_maxent_10km"), pattern = "\\.asc$", full.names = TRUE) %>%
    grep("bio2_1981-2010_global.asc|bio11_1981-2010_global.asc|bio12_1981-2010_global.asc|bio15_1981-2010_global.asc", ., value = TRUE)

```

I will create rasters of the environmental covariates, stack and gather summary statistics. I will also shorten their names and exclude possible operators from layer names (for example, using the dash symbol was found to interfere with `SDMtune` making predictions for tables downstream).

```{r naming object}

# layer name object. Check order of layers first
env_layer_names <- c("bio11", "bio12", "bio15", "bio2")

```

```{r stack historical covariates and make naming consistent, message = FALSE}

# global rasters
# stack env covariates
x_global_hist_env_covariates <- terra::rast(x = x_global_hist_env_covariates_list)

# attributes
nlyr(x_global_hist_env_covariates)
names(x_global_hist_env_covariates)
minmax(x_global_hist_env_covariates)
# ext(x_global_hist_env_covariates)
# crs(x_global_hist_env_covariates)

# I will change the name of the variables because they are throwing errors in SDMtune
names(x_global_hist_env_covariates) <- env_layer_names
# confirmed- SDMtune doesnt like dashes in column names (it is read as a mathematical operation)


```

```{r remove lists}

rm(x_global_hist_env_covariates_list)

```

```{r prepare SWD test objects}

# invaded asian
regional_invaded_asian_test_SWD <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_global_hist_env_covariates, 
  p = regional_invaded_asian_test, 
  a = a_regional_invaded_asian_background_points,
  verbose = TRUE # print helpful messages
  )

# regional native
regional_native_test_SWD <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_global_hist_env_covariates,
  p = regional_native_test, 
  a = a_regional_native_background_points,
  verbose = TRUE # print helpful messages
  )

# regional invaded_n_american
regional_invaded_n_american_test_SWD <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_global_hist_env_covariates, 
  p = regional_invaded_n_american_test, 
  a = a_regional_invaded_n_american_background_points,
  verbose = TRUE # print helpful messages
  )

```

## Plot

Next, I will use `SDMtune` to build the ROC plots.

```{r plot ROC per model}

regional_native_ROC <- SDMtune::plotROC(
  model = regional_native_model,
  test = regional_native_test_SWD
) %>%
  ggplot_build()

regional_invaded_n_american_ROC <- SDMtune::plotROC(
  model = regional_invaded_n_american_model,
  test = regional_invaded_n_american_test_SWD
) %>%
  ggplot_build()

regional_invaded_asian_ROC <- SDMtune::plotROC(
  model = regional_invaded_asian_model,
  test = regional_invaded_asian_test_SWD
) %>%
  ggplot_build()
           
```

```{r ROC plot- regional ensemble}

ROC_ensemble <- ggplot() +
  # native model data
  geom_line(data = regional_native_ROC$data[[1]], aes(x = x, y = y, color = "Rn (native)", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_n_american model data
  geom_line(data = regional_invaded_n_american_ROC$data[[1]], aes(x = x, y = y, color = "Ri.NAmerica", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_asian model data
  geom_line(data = regional_invaded_asian_ROC$data[[1]], aes(x = x, y = y, color = "Ri.Asia", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # midpoint line
  geom_abline(slope = 1, linetype = "dashed") +
  # scales
  scale_x_continuous(name = "False positive rate (1 - specificity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  scale_y_continuous(name = "True positive rate (sensitivity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  labs(
    title = "ROC curve (true vs false positive rate) for 'regional_ensemble' models"
    ) +
  theme_bw() +
  # aes
  scale_color_manual(
    name = "model",
    values = ensemble_colors,
    aesthetics = "color"
  ) +
  scale_linetype_manual(
    name = "Data segment",
    values = c("solid", "dotted"),
    labels = c("Test", "Train"),
    guide = guide_legend(reverse = TRUE)
  ) +
  theme(legend.position = "bottom") +
  coord_fixed(ratio = 1)

ROC_ensemble

```

```{r save plot- AUC, eval = FALSE, include = FALSE}

ggsave(
  ROC_ensemble, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "ROC_plot_regional_ensemble.jpg"
    ),
  height = 8, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```

We see that the regional-scale model has the highest testing AUC (0.99), followed by the invaded Asian model (0.775) and finally the invaded North American model (0.543). The invaded North American model is only slightly better than random (0.5), which is consistent with its low TSS score (0.32). Thankfully, however, we have significantly downweighted this model in our ensemble, so it should not have a large impact on our final risk map.

# Add global model to ROC plot

Now that I have plotted a separate figure for the three regional_ensemble models, I will create an aggregated figure that includes the ROC plot for all three regional_ensemble models and the global model. This will allow for a direct comparison of the AUC (goodness of fit) between the regional_ensemble and global models.

## ROC

```{r set wd 1.5}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

```{r load data objects, message = FALSE}

# model
global_model <- read_rds(file = file.path(mypath, "slf_global_v5", "global_model.rds"))
# background csv
global_background <- read_csv(file = file.path(here::here(), "vignette-outputs", "data-tables", "global_background_points_v5.csv"))
# all 
slf_points <- readr::read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2026-07-30.rds")) %>%
  dplyr::select(-species) 

ensemble_colors <- c(
  "Rn (native)" = "#4daf4a",
  "Ri.NAmerica" =  "#e41a1c",
  "Ri.Asia" = "#377eb8",
  "global" = "black"
)

```

```{r prepare global SWD and separate test obj}

global_SWD <- SDMtune::prepareSWD(
  species = "Lycorma delicatula",
  env = x_global_hist_env_covariates,
  p = slf_points,
  a = global_background,
  verbose = TRUE # print helpful messages
  )

# split into train and test sets
global_trainTest <-  SDMtune::trainValTest(
  x = global_SWD,
  test = 0.2, # set used for predicting maps and suitability values
  only_presence = TRUE, # only split presences and recycle bg points
  seed = 3 # should be the same as the OG because same seed
)

# separate off testing data
global_test_SWD <- global_trainTest[[2]]

```


```{r add global ROC}

global_ROC <- SDMtune::plotROC(
  model = global_model@models[[5]], # just pick one because they did not vary too much
  test = global_test_SWD
) %>%
  ggplot_build()

```

```{r ROC plot- global added}

ROC_ensemble_global <- ggplot() +
  # global
  geom_line(data = global_ROC$data[[1]], aes(x = x, y = y, color = "global", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # native model data
  geom_line(data = regional_native_ROC$data[[1]], aes(x = x, y = y, color = "Rn (native)", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_n_american model data
  geom_line(data = regional_invaded_n_american_ROC$data[[1]], aes(x = x, y = y, color = "Ri.NAmerica", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # invaded_asian model data
  geom_line(data = regional_invaded_asian_ROC$data[[1]], aes(x = x, y = y, color = "Ri.Asia", group = group, linetype = as.factor(group)), linewidth = 0.8) +
  # midpoint line
  geom_abline(slope = 1, linetype = "dashed") +
  # scales
  scale_x_continuous(name = "False positive rate (1 - specificity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  scale_y_continuous(name = "True positive rate (sensitivity)", breaks = c(0, 0.25, 0.5, 0.75, 1), limits = c(0, 1)) +
  labs(
    title = "ROC curve (true vs false positive rate) for 'regional_ensemble' and 'global' models"
    ) +
  theme_bw() +
  # aes
  scale_color_manual(
    name = "model",
    values = ensemble_colors,
    aesthetics = "color"
  ) +
  scale_linetype_manual(
    name = "Data segment",
    values = c("solid", "dotted"),
    labels = c("Test", "Train"),
    guide = guide_legend(reverse = TRUE)
  ) +
  theme(legend.position = "bottom") +
  coord_fixed(ratio = 1)

ROC_ensemble_global

```

```{r save plot- AUC combined, eval = FALSE, include = FALSE}

ggsave(
  ROC_ensemble_global, 
  filename = file.path(
    here::here(), "vignette-outputs", "figures", "ROC_plot_regional_ensemble_global.jpg"
    ),
  height = 8, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```

# Confusion Matrix- Sensitivity and Specificity of Models

I will create a table for some important summary statistics for the models. These include sensitivity, specificity, commission error, and omission error. These are calculated according to the cloglog MTSS threshold per model.

```{r set wd 2}

mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

```

We calculate these according to the cloglog MTSS threshold per model

```{r import confusion matrices, message = FALSE, results = "hide"}

# global
global_conf_matr <- read_csv(file = file.path(mypath, "slf_global_v5", "global_thresh_confusion_matrix_all_iterations.csv")) %>%
  # only MTSS threshold
  slice(7)


# regional models
# native
regional_native_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_native_v5", "regional_native_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_asia
regional_invaded_asian_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_asian_v4", "regional_invaded_asian_thresh_confusion_matrix.csv")) %>%
  slice(7)
# invaded_n_american_NAmerica
regional_invaded_n_american_conf_matr <- read_csv(file = file.path(mypath, "slf_regional_invaded_n_american_v9", "regional_invaded_n_american_thresh_confusion_matrix.csv")) %>%
  slice(7)

```

```{r display plot}

global_conf_matr

```


I will create some input datasets which contain the number of test positives per model.

```{r input datasets}

global_test_pos <- sum(global_test_SWD@pa == 1)
# regional
regional_native_test_pos <- nrow(regional_native_test)
regional_invaded_n_american_test_pos <- nrow(regional_invaded_n_american_test)
regional_invaded_asian_test_pos <- nrow(regional_invaded_asian_test)

```

```{r calculate sens and spec}

model_sens_spec <- tibble(
  metric = rep(c("sensitivity", "specificity", "commission error", "omission error"), 4),
  description = rep(c("true_positive rate", "true_negative rate", "1 - true_negative rate", "false_negative rate"), 4),
    model = c(
    rep("global", 4),
    rep("regional native", 4),
    rep("regional invaded North American", 4),
    rep("regional invaded Asian", 4)
  ),
  # cloglog MTSS threshold
  thresh = "MTSS.cloglog",
  thresh_value = c(
    rep(global_conf_matr$threshold_value, 4), 
    rep(regional_native_conf_matr$threshold_value, 4), 
    rep(regional_invaded_n_american_conf_matr$threshold_value, 4),
    rep(regional_invaded_asian_conf_matr$threshold_value, 4)
    ),
  # the number of test positives
  test_pos = c(
    rep(global_test_pos, 4), # global, this counts the number of positives in the model (the 1s in the training data)
    rep(regional_native_test_pos, 4),  # native
    rep(regional_invaded_n_american_test_pos, 4), # invaded_NAmerica
    rep(regional_invaded_asian_test_pos, 4) # invaded_Asia
  ),
  # the number of background points
  bg_neg = c(
    rep(20000, 4), # global
    rep(10000, 12) # regional
  ),
  value = c(
    # global
    round(global_conf_matr$tp_mean / global_test_pos, 4), # sensitivity (the tp rate)
    round(global_conf_matr$tn_mean / 20000, 4), # specificity (the tn rate)
    round(1 - (global_conf_matr$tn_mean / 20000), 4),  # commission error (1 - tn rate)
    round((global_conf_matr$fn_mean / 20000), 4), # omission error (the fn rate)
    # native
    round(regional_native_conf_matr$tp / regional_native_test_pos, 4), 
    round(regional_native_conf_matr$tn / 10000, 4), 
    round(1 - (regional_native_conf_matr$tn / 10000), 4), 
    round((regional_native_conf_matr$fn / 10000), 4),
    # invaded_NAmerica
    round(regional_invaded_n_american_conf_matr$tp / regional_invaded_n_american_test_pos, 4), 
    round(regional_invaded_n_american_conf_matr$tn / 10000, 4), 
    round(1 - (regional_invaded_n_american_conf_matr$tn / 10000), 4), 
    round((regional_invaded_n_american_conf_matr$fn / 10000), 4),
    # invaded_asia
    round(regional_invaded_asian_conf_matr$tp / regional_invaded_asian_test_pos, 4), 
    round(regional_invaded_asian_conf_matr$tn / 10000, 4), 
    round(1 - (regional_invaded_asian_conf_matr$tn / 10000), 4), 
    round((regional_invaded_asian_conf_matr$fn / 10000), 4)
  )
)

```

```{r save as .csv, eval = FALSE, include = FALSE}

write_csv(model_sens_spec, file.path(here::here(), "vignette-outputs", "data-tables", "model_sensitivity_specificity.csv"))

```

I will also write it to a kable so it can be printed

```{r write and display sens spec kable}

# add column formatting
model_sens_spec <- dplyr::mutate(model_sens_spec, bg_neg = scales::label_comma() (bg_neg))

# convert to kable
model_sens_spec_kable <- knitr::kable(x = model_sens_spec, format = "html", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)

model_sens_spec_kable

```

```{r save tables from above if desired, eval = FALSE, include = FALSE}

# save as .html
kableExtra::save_kable(
  model_sens_spec_kable, 
  file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
  self_contained = TRUE,
  bs_theme = "simplex"
  )

# initialize webshot by 
# webshot::install_phantomjs()
# convert to pdf
webshot::webshot(
  url = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"),
  file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.jpg"),
  zoom = 2
)

# file.remove(file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.html"))

# convert to kable
model_sens_spec_kable_LX <- knitr::kable(x = model_sens_spec, format = "latex", escape = FALSE) %>%
  kableExtra::kable_styling(bootstrap_options = "striped", full_width = TRUE)
# save as latex
kableExtra::save_kable(
  model_sens_spec_kable_LX, 
  file = file.path(here::here(), "vignette-outputs", "figures", "model_sensitivity_specificity.tex"),
  self_contained = TRUE,
  bs_theme = "simplex"
  )

```

We see once again that all models have adequate sensitivity and specificity, but the outlier is the invaded North American model. This model actually had a sensitivity of 0, which means it did not detect any of its test points. The testing range for this model was the invaded range in Asia, where the model predicted no area of suitability above the MTSS threshold. This is likely due to the significant difference in precipitation seasonality. The current invaded North American range has a very narrow range of precipitation seasonality, as compared with the test area in Korea and Japan, which have seasonal monsoons. You can observe this by looking at the variable curve for bio 15; the invaded asian and North American models occupied entirely different ranges of precipitation seasonality. Thus, the training range likely did not translate well to the testing range in Asia.

Otherwise, all models have an adequate fit. The ensemble accounts for the bad fit of the North American model, which we will continue to include due to the potential for this range to expand over time and the important data it includes.

# References

1. Allouche, O., Tsoar, A., & Kadmon, R. (2006). Assessing the accuracy of species distribution models: Prevalence, kappa and the true skill statistic (TSS). Journal of Applied Ecology, 43(6), 1223–1232. https://doi.org/10.1111/j.1365-2664.2006.01214.x

2. Jiménez‐Valverde, A. (2012). Insights into the area under the receiver operating characteristic curve (AUC) as a discrimination measure in species distribution modelling. Global Ecology and Biogeography, 21(4), 498–507. https://doi.org/10.1111/j.1466-8238.2011.00683.x

3. Zou, K. H., O’Malley, A. J., & Mauri, L. (2007). Receiver-Operating Characteristic Analysis for Evaluating Diagnostic Tests and Predictive Models. Circulation, 115(5), 654–657. https://doi.org/10.1161/CIRCULATIONAHA.105.594929
