---
title: "Example usage of create_risk_report(), applied to create reports for French Viticultural areas"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu, [ORCID](https://orcid.org/0009-0001-2338-7928)]"
date: "2025-08-07"
---

# Overview

In the last few vignettes, I have created various outputs that help me to analyze the risk of SLF establishment at for global viticultural regions. I created risk maps and tables, range shift maps and tables, risk quadrant maps for viticultural regions and SLF populations, and have calculated various summary statistics such as omission and commission error, area under the curve, and the optimal suitability threshold for each model. 

In this vignette, I will use my function `create_risk_report()` to refine each of these outputs (excluding the summary statistics, which area model-specific) for countries and provinces of key importance for global viticulture. These outputs will crop maps and viticultural region risk assessments for each region of interest. I will begin by creating a report for each country in our input the `wineries_esri54017_tidied` dataset. I will also create reports for key areas where viticulture is expected to expand or decline under climate change. This vignette will also outline the best practice for usage of my function `create_risk_report()`.

# Setup

```{r library packages, message = FALSE}

# general tools
library(tidyverse)
library(cli)
library(here)
library(common)

# spatial data
library(terra)
library(sf)
library(rnaturalearth)
library(rnaturalearthhires)
library(rnaturalearthdata)
library(ggspatial)

# aesthetics
library(kableExtra)
library(formattable)
library(webshot)
library(webshot2)
library(ggnewscale) ## MUST use version 0.4.10- 0.5.0 has a bug that breaks this function

# for analysis
library(patchwork)
library(gginnards)

library(scari)

```

# Create internal dataset for function

You should only need to do this once, and again each time you change the input data.

I will start by creating an internal dataset to be used by the function. These datasets were created in the previous vignettes and are stored in the `data` or `data-raw` folders of the package. I will import these datasets and save them as internal datasets for the function to use.

If you have other datasets to use, you can replace the datasets below with your own, so long as they are in a similar format.

```{r datasets for function data import, eval = FALSE}

# .csv files
threshold_exponential_values <- read.csv(file = file.path(here::here(), "data-raw", "threshold_exponential_values.csv"))
# .rds files
IVR_locations <- readr::read_rds(file = file.path(here::here(), "data", "wineries_esri54017_tidied.rds"))
summary_global <- readr::read_rds(file = file.path(here::here(), "data", "global_threshold_values.rds"))
summary_regional_ensemble <- readr::read_rds(file = file.path(here::here(), "data", "ensemble_threshold_values.rds"))
# predicted xy suitability
xy_global_hist <- readr::read_rds(file = file.path(here::here(), "data", "global_wineries_1981-2010_xy_pred_suit.rds"))
xy_global_future <- readr::read_rds(file = file.path(here::here(), "data", "global_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds"))
xy_regional_ensemble_hist <- readr::read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_1981-2010_xy_pred_suit.rds"))
xy_regional_ensemble_future <- readr::read_rds(file = file.path(here::here(), "data", "regional_ensemble_wineries_2041-2070_GFDL_ssp_mean_xy_pred_suit.rds"))

```

I need to make some edits to the xy datasets.

```{r rename columns, eval = FALSE}

xy_global_hist <- dplyr::rename(xy_global_hist, "xy_global_hist" = "xy_global_1995")
xy_global_future <- dplyr::rename(xy_global_future, "xy_global_future" = "xy_global_2055")
xy_regional_ensemble_hist <- dplyr::rename(xy_regional_ensemble_hist, "xy_regional_ensemble_hist" = "xy_regional_ensemble_1995")
xy_regional_ensemble_future <- dplyr::rename(xy_regional_ensemble_future, "xy_regional_ensemble_future" = "xy_regional_ensemble_2055")

```


```{r import raster files and save to new location for create_risk_report function, eval = FALSE}

# mypath
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models")

# import raster files
## historical raster
slf_binarized_hist <- terra::rast(x = file.path(mypath, "working_dir", "slf_binarized_summed_1981-2010.asc")) 
slf_binarized_future <- terra::rast(x = file.path(mypath, "working_dir", "slf_binarized_summed_2041-2070_ssp_mean_GFDL.asc")) 
slf_range_shift <- terra::rast(x = file.path(mypath, "working_dir", "slf_range_shift_summed_ssp_mean_GFDL.asc"))

# write to proper folder
terra::writeRaster(
  slf_binarized_hist, 
  filename = file.path(here::here(), "vignette-outputs", "rasters", "slf_binarized_summed_1981-2010.asc"), 
  overwrite = FALSE
)

terra::writeRaster(
  slf_binarized_future, 
  filename = file.path(here::here(), "vignette-outputs", "rasters", "slf_binarized_summed_2041-2070_ssp_mean_GFDL.asc"), 
  overwrite = FALSE
)

terra::writeRaster(
  slf_range_shift, 
  filename = file.path(here::here(), "vignette-outputs", "rasters", "slf_range_shift_summed_ssp_mean_GFDL.asc"), 
  overwrite = FALSE
)


```

```{r internal data object, eval = FALSE}

# transform all data frames to internal dataset
usethis::use_data(
  threshold_exponential_values, IVR_locations, summary_global, summary_regional_ensemble, xy_global_hist, xy_global_future, xy_regional_ensemble_hist, xy_regional_ensemble_future,
  internal = TRUE,
  overwrite = FALSE
)

```

The internal dataset `sysdata.rda` is used for this function in the remainder of the vignette; this file is stored at `root/R`.

# Example usage for `create_risk_report()`: France

I will now pull out a specific example of France to analyze and demonstrate usage of this function. I will use France as a case study. Here is an example of the code we could use to retrieve a report for France. We will work through each argument,

The first thing I will need to do is to search for the correct [ISO 3166-1 country code](https://en.wikipedia.org/wiki/ISO_3166-1_alpha-3) corresponding to our country of interest. France is "FRA", so I will enter that in `locality.iso`. I can also enter the name of the country for reporting purposes, but it is not required for countries (only states/provinces).

```{r generate a report for France, message = FALSE, results = "hide"}

scari::create_risk_report(
  locality.iso = "fra", # A3-iso
  locality.name = "France", #name
  locality.type = "country", # record type
  # saving output
  mypath = file.path(here::here(), "vignette-outputs", "reports", "France"), # the output location
  create.dir = FALSE, # not saving so not necessary
  save.report = FALSE, # dont save, this is an example
  raster.path = file.path(here::here(), "vignette-outputs", "rasters"), # path to rasters
  # aesthetics- dont necessarily need to be specified
  buffer.dist = 20000, # 20km, this is the distance at which buffers should be drawn on maps to indicate IVR suitability predictions
  period.present = "1981-2010",
  period.projected = "2041-2070",
  model.projected = "GFDL-ESM4",
  ssp.projected = "ssp_126_370_585",
  crs = "ESRI:54017"
)

```

First, we need to look up the A3-iso code for France. It is "FRA". We also need to specify that this is a country.

Now, we need to specify arguments for the report output. `mypath` specifies the path to the directory where this will be saved. It is not necessary if the report will not be saved (ie, `create.dir = FALSE` and `save.report = FALSE`). The argument `raster.path` specifies where the rasters are stored for this function. They are pre-loaded above and by default are stored in `root/vignette-outputs/rasters`.

Next, we can optionally specify arguments for the aesthetics. If not specified, they will default to a specified value. I have listed the defaults in the example. *Note* that if you change any of these from their defaults, you should ensure that you have first changed the input datasets, which I created above. This would require you to rerun the package vignettes. I will give an explanation and list out the arguments below:

`buffer.dist`: this specifies the distance at which a buffer is drawn around important viticultural regions (IVRs) to calculate the suitability. The max suitability is taken from this buffer region. If left blank, it will assume that you instead used a simple point-wise calculation of suitability (aka, the suitability was taken at the exact point locations of IVRs). *NOTE*: if you choose to change this here, it will only change the aesthetic and not the calculations (the data are created in vignette 130 and pre-loaded above.) You should change the buffer distance used in that vignette if you wish to make this change. 

`period.present`: The time period of the historical data.

`period.projected`: The time period for the projected future data.

`model.projected`: The model used for the projected future data.

`ssp.projected`: The shared socioeconomic pathway (SSP) used for the projected future data. This is a scenario of future climate change that is used to project future conditions.

`crs`: The crs chosen for all input data and rasters. Note that changing this would require a complete rerun of the package vignettes.

Finally, we have `map.style`, which is a purely aesthetic argument. It should be specified as a list argument if not the default (the default is shown below). Again, this does not need to be specified.

```{r default of map.style, eval = FALSE}

  # map stype argument as a list
  map.style <- list(
      xlab("UTM_eastings"), # if raster is in lonlat, label as lon/lat, otherwise UTM
      ylab("UTM_northings"), # if raster is in lonlat, label as lon/lat, otherwise UTM
      # aesthetics
      theme_classic(),
      theme(
        # legend
        legend.position = "bottom",
        legend.key = element_rect(color = "black")
      ),
      guides(fill = guide_legend(nrow = 1, byrow = TRUE)),
      # scales
      scale_x_continuous(expand = c(0, 0)),
      scale_y_continuous(expand = c(0, 0))
    )

```

Now, lets call the various elements of this report.

First, we have the summary of this report, which shows important data.

```{r analyze France report 1}

france_L_delicatula_risk_report[["Report_info"]]

```

Next, we have a summary of the viticultural regions in France, which shows each IVR (winery), its published coordinates, and the risk associated with each winery:

```{r analyze France report 2}

france_L_delicatula_risk_report[["viticultural_regions_list"]]

```

Here are some of the column explanations:

- columns that include the word "risk" are a calculation of the risk level, on a scale from 0-1. There are columns for both the present and future projections of risk. 
- Columns labeled "risk_level" and the current level of risk; low, moderate, high, or extreme.
- the risk_shift column shows the present and future risk, separated by an underscore
- the risk_shift_count column tallys how many risk levels were gained or lost, in order of the risk level.

Now, we will call up the risk maps, which depict the level of risk and the location of any wineries.

```{r analyze France report 3, warning = FALSE}

france_L_delicatula_risk_report[["risk_maps"]][["present_risk_map"]]

france_L_delicatula_risk_report[["risk_maps"]][["future_risk_map"]]

```

Unfortunately, the map did not format well because France possesses some territories outside its mainland. We can fix this with a few simple lines of code; we merely need to edit the axes of the plot:

```{r change view on maps 1, warning = FALSE}

france_current <- france_L_delicatula_risk_report[["risk_maps"]][["present_risk_map"]] +
  xlim(-482431.4012544825, 964862.802508965) +
  ylim(4804640.544946612, 5695846.383707797)

france_current

```

```{r change view on maps 2, warning = FALSE}

france_future <- france_L_delicatula_risk_report[["risk_maps"]][["future_risk_map"]] +
  xlim(-482431.4012544825, 964862.802508965) +
  ylim(4804640.544946612, 5695846.383707797)

france_future

```

Now we can see the map of the mainland much better.

The next table shows the proportion of the map that is occupied by each risk level. This is a useful table to see how much of the viticultural area is at risk.

```{r analyze France report 4}

france_L_delicatula_risk_report[["risk_maps_prop_area_table"]]

```

Next, we have the viticultural risk plot, which shows the risk level of each winery in France. This is a useful plot to see the overall trend of how the wineries in France are changing risk over time. The arrows point from a vineyard in the present, to its risk in the future.

```{r analyze France report 5}

france_L_delicatula_risk_report[["viticultural_risk_plot"]]

```

Next is the viticultural risk table, which shows a tally of the wineries at each risk level, both now and in the future. This is a useful table to see how many wineries are at each risk level, and how that changes over time. It serves to quantify the risk plot above. Negative numbers are decreasing in risk over time, and positive numbers are increasing in risk over time.

```{r analyze France report 6}

france_L_delicatula_risk_report[["viticultural_risk_table"]]

```

Next, we have a map of the shift in risk for SLF over time. Green areas will gain risk in the future, while blue areas will lose risk in the future. This is a useful map to see where SLF is expected to shift its range over time. The white areas remain consistently at risk over time, while the grey areas are never at risk.

```{r analyze France report 7, warning = FALSE}

france_L_delicatula_risk_report[["range_shift_map"]]

```

We cannot see this map well either; lets change the view on it.

```{r change view on maps 3, warning = FALSE}

france_shift <- france_L_delicatula_risk_report[["range_shift_map"]] +
  xlim(-482431.4012544825, 964862.802508965) +
  ylim(4804640.544946612, 5695846.383707797)

france_shift

```

Finally, we have a table which calculates the total area for each of the categories in the above map.

```{r analyze France report 8}

france_L_delicatula_risk_report[["range_shift_table"]]

```

## Example of subsetting reports- Bordeaux region of France

Looking back at the list of important viticultural regions we retrieved above, we can see that France has 70 total results for which we produced a report. It can be helpful to see the overall shift in risk due to climate change at the scale of the country, but you might find it difficult to discern particular regions of interest at this scale. One option you have is to refine the results at the state/provincial scale. To do this, simply set `locality.type = "state_province"` and enter the name of your province of interest into the `locality.name` argument. 

We can see in the above report that there are many entries for a region referred to as "burgundy_bourgogne_burgundy" (of course, this includes some duplicate naming and multiple namings- but the data are still there). We will retrieve entries only for this region of interest. The `Region` column represents states and provinces (admin-1 groups). To retrieve records for this region, we will need the [ISO 3166-2 state/province code](https://en.wikipedia.org/wiki/ISO_3166-2) which corresponds to this region. If we click this link and France on the "Entry" column of the table, we find that "Burgundy" corresponds to the region of "Burgundy-Franche-Comté", which has the following iso code: "FR-BFC". We will enter the code into `locality.iso.2` and the name into `locality.name`.

```{r generate a report for France, message = FALSE, results = "hide"}

scari::create_risk_report(
  locality.iso = "fra", # A3-iso
  locality.iso.2 = "FR-BFC",
  locality.name = "Burgundy-Franche-Comté", #name
  locality.type = "state_province", # record type
  # saving output
  mypath = file.path(here::here(), "vignette-outputs", "reports", "France"), # the output location
  create.dir = FALSE, # not saving so not necessary
  save.report = FALSE, # dont save, this is an example
  raster.path = file.path(here::here(), "vignette-outputs", "rasters"), # path to rasters
  # aesthetics- dont necessarily need to be specified
  buffer.dist = 20000, # 20km, this is the distance at which buffers should be drawn on maps to indicate IVR suitability predictions
  period.present = "1981-2010",
  period.projected = "2041-2070",
  model.projected = "GFDL-ESM4",
  ssp.projected = "ssp_126_370_585",
  crs = "ESRI:54017"
)

```

Let's take a look at the number of results now:

```{r analyze Bordeaux report 1}

burgundy_franche_comté_L_delicatula_risk_report[["viticultural_regions_list"]]

```

Now analyze risk plot and table:

```{r analyze Bordeaux report 2}

burgundy_franche_comté_L_delicatula_risk_report[["viticultural_risk_plot"]]

```

```{r analyze Bordeaux report 3}

burgundy_franche_comté_L_delicatula_risk_report[["viticultural_risk_table"]]

```

From this analysis, we can see now that IVR's in the "Burgundy" region of France are all at extreme risk (suitable in both models) of SLF establishment presently, and this will not change under climate change.

# Apply function to all countries containing IVRs

## loop to apply function to wineries dataset

Primarily, I would like to use the wineries dataset to create a report for each unique country on the list. Our winery records contain 64 unique countries that I will create reports for. These reports will be saved to `vignette-outputs/reports`. 

```{r get country iso codes}

# import wineries dataset
IVR_locations <- readr::read_rds(file = file.path(here::here(), "data", "wineries_esri54017_tidied.rds"))


n_distinct(IVR_locations$Country)
# there are 64 unique countries on the list

# get iso codes using rnaturalearth data
admin_0_countries <- rnaturalearth::ne_load(
      scale = 10,
      type = "admin_0_countries",
      category = "cultural",
      destdir = file.path(here::here(), "data-raw", "ne_countries"),
      returnclass = "sf"
    ) %>%
  dplyr::select(ADMIN, ADM0_A3)

# replace names in sf
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "United States of America", "United States")
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "South Korea", "Republic of Korea")
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "Palestine", "Palestinian territories")
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "Cabo Verde", "Cape Verde")
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "Czechia", "Czech Republic")
admin_0_countries$ADMIN <- stringr::str_replace_all(admin_0_countries$ADMIN, "Republic of Serbia", "Serbia")

# correct misspelling
IVR_locations$Country <- stringr::str_replace_all(IVR_locations$Country, "Algiera", "Algeria")

# join
IVR_locations_iso <- left_join(IVR_locations, admin_0_countries, join_by(Country == ADMIN)) %>%
  dplyr::rename("iso_a3" = "ADM0_A3")
  
  
# check 
n_distinct(IVR_locations_iso$iso_a3)
# it was successful, 64 unique iso codes

# sort dataset
unique_iso <- sort(unique(IVR_locations_iso$iso_a3))
# paired country names
unique_name <- IVR_locations_iso %>%
  group_by(iso_a3) %>%
  summarize(country = unique(Country)) %>%
  dplyr::select(country) %>%
  as.matrix() %>%
  as.character()

```

```{r test loop, eval = FALSE}

for (i in seq_along(unique_iso)){
  print(c(unique_name[i],unique_iso[i]))
}


length(unique_name)

```

We ran checks and everything seems correct. Above is a list of countries that will get a report, 64 in total. Now, I will run a loop to create those reports and save the, to `root/vignette-outputs/reports`.

```{r loop for report, eval = FALSE, warning = FALSE, message = FALSE, results = "hide"}

# loop thru country iso codes and names together
for (a in seq_along(unique_iso)) {
  
  scari::create_risk_report(
    locality.iso = unique_iso[a], # iso codes
    locality.name = unique_name[a], # names
    locality.type = "country",
    mypath = file.path(here::here(), "vignette-outputs", "reports", unique_name[a]),
    create.dir = TRUE, # create subdirectory for plots
    save.report = TRUE,
    buffer.dist = 20000 # distance at which buffers should be drawn on maps to indicate IVR suitability predictions
    )
  
}

```

Some of these maps might need to be edited to zoom in on the countries, because some countries have geographically separated territories that are not relevant to the analysis. The code to zoom in on those plots will be in the following section.

```{r unique name loop, eval = FALSE, include = FALSE}

# deprecated loop to save reports 

unique_name_loop <- unique_name %>%
    tolower() %>%
    stringr::str_replace_all(., pattern = " |-", replacement = "_") %>%
    stringr::str_remove_all(., pattern = "['/.()]") # all these symbols

for (i in seq_along(unique_name_loop)) {
  
    readr::write_rds(get(paste0(unique_name_loop[i], "_L_delicatula_risk_report")), file = file.path(here::here(), "vignette-outputs", "reports", unique_name[i], paste0(unique_name_loop[i], "_L_delicatula_risk_report.rds")))
                                                                                     
}

```
