---
title: "Create risk reports specifically for the USA and US States"
output: rmarkdown::html_document
author: 
    - "Samuel M. Owens^[Temple University\n[Center for Biodiversity](http://www.biodiversitycenter.org/)\n[iEco Lab](https://www.iecolab.org/)\nsam.owens@temple.edu]"
date: "2025-08-07"
---

VIGNETTE UNDER CONSTRUCTION

```{r library packages, message = FALSE}

# general tools
library(tidyverse)
library(cli)
library(here)
library(common)

# spatial data
library(terra)
library(sf)
library(rnaturalearth)
library(rnaturalearthhires)
library(rnaturalearthdata)
library(ggspatial)

# aesthetics
library(kableExtra)
library(formattable)
library(webshot)
library(webshot2)
library(ggnewscale) ## MUST use version 0.4.10- 0.5.0 has a bug that breaks this function

# for analysis
library(patchwork)
library(gginnards)

library(scari)

```

**Note:** I will be setting the global options of this document so that only certain code chunks are rendered in the final .html file. I will set the `eval = FALSE` so that none of the code is re-run (preventing files from being overwritten during knitting) and will simply overwrite this in chunks with plots.

```{r set global chunk opts, include = FALSE}

# eval = FALSE so that outputs not overwritten
knitr::opts_chunk$set(eval = FALSE)

```

## USA

USA as a country

```{r USA reports}

scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "united states",
  locality.type = "country",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

```

```{r edit USA reports, eval = FALSE}

### USA
# the USA map usually comes including Alaska and hawaii, which we do not want

USA_current <- united_states_slf_risk_report[["risk_maps"]][["present_risk_map"]] +
  xlim(-12060785.031362064, -6271608.216308273) + # the boundaries in UTM
  ylim(3091555.5611515422, 5614050.1030608015)

USA_future <- united_states_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  xlim(-12060785.031362064, -6271608.216308273) +
  ylim(3091555.5611515422, 5614050.1030608015)

USA_shift <- united_states_slf_risk_report[["range_shift_map"]] +
  xlim(-12060785.031362064, -6271608.216308273) +
  ylim(3091555.5611515422, 5614050.1030608015)


#save
ggsave(
  USA_current,
  filename = file.path(here::here(), "vignette-outputs", "reports", "United States", "united_states_risk_map_present_edited.jpg"),
  height = 8,
  width = 10,
  device = jpeg,
  dpi = "retina"
)

ggsave(
  USA_future,
  filename = file.path(here::here(), "vignette-outputs", "reports", "United States", "united_states_risk_map_future_edited.jpg"),
  height = 8,
  width = 10,
  device = jpeg,
  dpi = "retina"
)

ggsave(
  USA_shift,
  filename = file.path(here::here(), "vignette-outputs", "reports", "United States", "united_states_shift_map_edited.jpg"),
  height = 8,
  width = 10,
  device = jpeg,
  dpi = "retina"
)

# save USA report for other uses
readr::write_rds(united_states_slf_risk_report, file = file.path(here::here(), "vignette-outputs", "reports", "United States", "united_states_slf_risk_report.rds"))

```

### US States

```{r CA and PA}

# try state level
scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "california",
  locality.type = "state_province",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "pennsylvania",
  locality.type = "state_province",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

```

# US states with projected increase in viticultural sutiability

The following areas may increase in suitability due to warming winter temperature minimums:

-   Michigan, USA
-   New Zealand

```{r states with projected increase in viticultural suitability}

# michigan
scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "michigan",
  locality.type = "state_province",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

# oregon
scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "oregon",
  locality.type = "state_province",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

# washington
scari::create_risk_report(
  locality.iso = "usa",
  locality.name = "washington",
  locality.type = "state_province",
  mypath = file.path(here::here(), "vignette-outputs", "reports", "United States"),
  create.dir = FALSE,
  save.report = FALSE,
  buffer.dist = 20000
)

```

```{r US states, eval = FALSE}

washington_risk_plot <- washington_slf_risk_report[["viticultural_risk_plot"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "A") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold")
    )

# edits
# remove geom_text
washington_risk_plot <- washington_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
washington_risk_plot$data[[4]]$size <- 1.5
washington_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#washington_risk_plot$data[[3]]$linewidth <- 0.1

washington_risk_plot <- ggplot_gtable(washington_risk_plot) %>%
  wrap_ggplot_grob()



oregon_risk_plot <- oregon_slf_risk_report[["viticultural_risk_plot"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "C") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold")
    )

oregon_risk_plot <- oregon_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
oregon_risk_plot$data[[4]]$size <- 1.5
oregon_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#oregon_risk_plot$data[[3]]$linewidth <- 0.1

oregon_risk_plot <- ggplot_gtable(oregon_risk_plot) %>%
  wrap_ggplot_grob()

  

michigan_risk_plot <- michigan_slf_risk_report[["viticultural_risk_plot"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "E") +
  theme(
    legend.position = "none", 
    panel.border = element_rect(size = 1, linetype = "solid", color = "black"), 
    plot.tag.position = c(0.2, 0.9),
    plot.tag = element_text(face = "bold")
    )

michigan_risk_plot <- michigan_risk_plot %>%
  gginnards::delete_layers(match_type = "GeomLabel") %>%
  ggplot_build()
# edit point size
michigan_risk_plot$data[[4]]$size <- 1.5
michigan_risk_plot$data[[5]]$size <- 1.5
# edit linewidth
#michigan_risk_plot$data[[3]]$linewidth <- 0.1

michigan_risk_plot <- ggplot_gtable(michigan_risk_plot) %>%
  wrap_ggplot_grob()


  # edit maps

washington_risk_map <- washington_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    ) +
  labs(tag = "B") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(-0.1, 0.9),
    plot.tag = element_text(face = "bold")
    )


washington_risk_map <- ggplot_build(washington_risk_map)
# edit point size
washington_risk_map$data[[3]]$size <- 1.5

washington_risk_map <- ggplot_gtable(washington_risk_map) %>%
  wrap_ggplot_grob()



oregon_risk_map <-  oregon_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    ) +
  labs(tag = "D") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(-0.1, 0.9),
    plot.tag = element_text(face = "bold")
    ) 

oregon_risk_map <- ggplot_build(oregon_risk_map)
# edit point size
oregon_risk_map$data[[3]]$size <- 1.5

oregon_risk_map <- ggplot_gtable(oregon_risk_map) %>%
  wrap_ggplot_grob()
 


michigan_risk_map <- michigan_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank(),
    axis.text = element_blank(),
    axis.ticks = element_blank()
    ) +
  labs(tag = "F") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(-0.1, 0.9),
    plot.tag = element_text(face = "bold")
    )
  
michigan_risk_map <- ggplot_build(michigan_risk_map)
# edit point size
michigan_risk_map$data[[3]]$size <- 1.5
# rebuild plot
michigan_risk_map <- ggplot_gtable(michigan_risk_map) %>%
  wrap_ggplot_grob()
  
  
```

```{r example states- patchwork, eval = FALSE}

example_states_patchwork <- (
  # washington top row
  washington_risk_plot + plot_spacer() + washington_risk_map + 
  # USA 2nd row
  oregon_risk_plot + plot_spacer() + oregon_risk_map + 
  # michigan 3rd row
  michigan_risk_plot + plot_spacer() + michigan_risk_map
  ) +
  # annotation
  plot_annotation(title = "Risk shift in US states of future interest for viticulture") +
  plot_layout(ncol = 3, nrow = 3, widths = unit(c(5.5, 1, 5.5), "cm"), heights = unit(5.5, "cm"))


```

```{r save plot- example states, eval = FALSE, include = FALSE}

ggsave(
  example_states_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "risk_shift_viticultural_states.jpg"),
  height = 8, 
  width = 10,
  device = jpeg,
  dpi = 500
  )

```

### Patchwork USA maps

```{r USA}

USA_risk_map_present <- united_states_slf_risk_report[["risk_maps"]][["present_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "Present") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(0.9, 0.20),
    plot.tag = element_text(face = "bold", size = 20)
    ) +
  xlim(-12060785.031362064, -6271608.216308273) + # the boundaries in UTM
  ylim(3091555.5611515422, 5614050.1030608015)

USA_risk_map_present <- ggplot_build(USA_risk_map_present)
# edit point size
USA_risk_map_present$data[[3]]$size <- 1.5

USA_risk_map_present <- ggplot_gtable(USA_risk_map_present) %>%
  wrap_ggplot_grob()
 


USA_risk_map_future <- united_states_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "Future") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(0.9, 0.2),
    plot.tag = element_text(face = "bold", size = 20)
    ) +
  xlim(-12060785.031362064, -6271608.216308273) + # the boundaries in UTM
  ylim(3091555.5611515422, 5614050.1030608015)

USA_risk_map_future <- ggplot_build(USA_risk_map_future)
# edit point size
USA_risk_map_future$data[[3]]$size <- 1.5

USA_risk_map_future <- ggplot_gtable(USA_risk_map_future) %>%
  wrap_ggplot_grob()
  
  
```

```{r USA- patchwork}

USA_patchwork <- (
  # romania top row
  USA_risk_map_present / USA_risk_map_future 
  ) +
  # annotation
  plot_annotation(title = "USA risk maps") 

USA_patchwork

```

```{r save plot- USA maps, eval = FALSE, include = FALSE}

ggsave(
  USA_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "USA_risk_maps.jpg"),
  height = 10, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```

### Patchwork washington maps

```{r washington, eval = FALSE}

WA_risk_map_present <-  washington_slf_risk_report[["risk_maps"]][["present_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "Present") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(0.9, 0.1),
    plot.tag = element_text(face = "bold", size = 20)
    ) 

WA_risk_map_present <- ggplot_build(WA_risk_map_present)
# edit point size
WA_risk_map_present$data[[3]]$size <- 1.5

WA_risk_map_present <- ggplot_gtable(WA_risk_map_present) %>%
  wrap_ggplot_grob()
 


WA_risk_map_future <-  washington_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  theme(
    plot.title = element_blank(),
    plot.subtitle = element_blank(),
    plot.caption = element_blank(),
    axis.title = element_blank()
    ) +
  labs(tag = "Future") +
  theme(
    legend.position = "none", 
    plot.tag.position = c(0.9, 0.1),
    plot.tag = element_text(face = "bold", size = 20)
    )

WA_risk_map_future <- ggplot_build(WA_risk_map_future)
# edit point size
WA_risk_map_future$data[[3]]$size <- 1.5

WA_risk_map_future <- ggplot_gtable(WA_risk_map_future) %>%
  wrap_ggplot_grob()
  
  
```

```{r patchwork- washington, eval = FALSE}

WA_patchwork <- (
  # romania top row
  WA_risk_map_present / WA_risk_map_future 
  ) +
  # annotation
  plot_annotation(title = "washington risk maps") 

```

```{r save plot- washington, eval = FALSE, include = FALSE}

ggsave(
  WA_patchwork, 
  filename = file.path(here::here(), "vignette-outputs", "figures", "washington_risk_maps.jpg"),
  height = 10, 
  width = 8,
  device = jpeg,
  dpi = "retina"
  )

```
