---
title: "Get started using scari"
output: rmarkdown::html_document
author:
    - "Samuel M. Owens^[Temple University, sam.owens@temple.edu, [ORCID](https://orcid.org/0009-0001-2338-7928)]"
date: "2026-06-23"
---

```{r library packages, echo = FALSE, message = FALSE}

# general tools
library(tidyverse)
library(cli)
library(here)
library(common)
library(devtools)

# spatial data
library(terra)
library(sf)

# visualization
library(kableExtra)

# specialized packages
library(rgbif)
library(kgc)

library(scari)

```

# How to use this project

## 1. Produce localized reports on SLF risk to viticulture

Our primary datasets, including risk maps and a viticultural risk analysis, can be accessed by executing the function `create_risk_report()`. This function creates a localized (at the country or state/provincial level) report of the risk for *Lycorma delicatula* to local viticulture. It can be applied to globally important wine growing regions because it uses our dataset `data/wineries_esri54017_tidied.rds`, which contains a sample (1,072) of the world's most important wine growing regions. Users can and should apply this function for their locality as we are projecting that the risk of *Lycorma delicatula* establishment will expand under climate change. *Lycorma delicatula* has been classified as being capable of (global) pan-invasion and viticulure has already been devastated in some regions, so there is high potential for viticultural damage under climate change (Huron et al, 2022).

In this vignette, I will give an example of how this function can be applied to both a country and state/province to assess risk at different scales. For a more complete example of the usage of the `create_risk_report()` function and all of its outputs, see the vignette [150_create_risk_report_example_usage](https://github.com/ieco-lab/scari/blob/master/vignettes/150_create_risk_report_example_usage.Rmd).

Our function recreates major data sets that we provide in our main analysis, including:

1. list of known important wine regions within the locality with predicted suitability values and levels
2. current and future risk maps for SLF establishment
3. range shift map of potential range expansion for L. delicatula under climate change
4. Viticultural quadrant plot depicting of the risk for SLF establishment for known wine regions within the locality. This plot depicts the intersection of our two modeled scales.
5. risk table quantifying the level of risk to vineyards according to the quadrant plot

Here is an example of a workflow for creating a localized risk map and how this might be applied. I will give an example of how I might interpret the risk for *Lycorma delicatula* establishment in the United States, and then I will show how to narrow its scope to a more localized region, Washington State. 

I might begin by creating a report at the national level. To create this, you will need the alpha-3 iso country code.

```{r generate report for USA, message = FALSE, warning = FALSE, results = "hide"}

scari::create_risk_report(
  locality.iso = "usa",
  locality.type = "country", # to specify the type of report
  buffer.dist = 20000, # in meters, this is the buffer zone defining the viticultural area
  create.dir = FALSE, # should a directory be created for the report? If TRUE, param mypath will be required
  save.report = FALSE # should the report be saved?
)

```

You should first note the three most important outputs of this function:

1. the risk maps
2. the accompanying viticultural risk quadrant plot
3. tabled list of viticultural regions in this locality

First, we observe the risk map, which depicts the risk of SLF establishment using the agreement between our regional-scale model ensemble and our global-scale model. More agreement on suitability mean higher risk.

```{r retrieve current risk map for USA, message = FALSE, warning = FALSE}

USA_current <- usa_slf_risk_report[["risk_maps"]][["present_risk_map"]] +
  # you may need to crop the map due to outlying territories, as we have done here
  xlim(-12060785.03, -6271608.22) + # the boundaries in UTM
  ylim(3091555.56, 5614050.10)

USA_current

```

**Fig 1.** Projected current risk of *Lycorma delicatula* establishment | USA

```{r retrieve projected future risk map for USA, message = FALSE, warning = FALSE}

USA_future <- usa_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  # you may need to crop the map due to outlying territories, as we have done here
  xlim(-12060785.03, -6271608.22) + # the boundaries in UTM
  ylim(3091555.56, 5614050.10)

USA_future

```

**Fig 2.** Projected future risk of *Lycorma delicatula* establishment under climate change | USA

The points on the map represent key viticultural regions. We have extracted the suitability of each viticultural region and depicted its quantitative shift in risk on our second output, the viticultural risk quadrant plot. This plot quantifies the level of risk along both modeled scales, both presently (light purple) and in the future (dark purple) under a predicted climate change shift (arrows):

```{r retrieve risk plot USA, echo = FALSE, message = FALSE, warning = FALSE}

USA_risk_plot <- usa_slf_risk_report[["viticultural_risk_plot"]] +
  labs(
    title = element_blank(),
    subtitle = element_blank(),
    caption = element_blank()
  ) +
  theme(legend.position = "none")

USA_risk_plot

```

**Fig. 3:** Projected shift in the risk for *Lycorma delicatula* establishment at key viticultural regions due to climate change | USA

The accompanying table provides a list of key viticultural regions and their geographical region (state/province), with predicted risk levels:

**Table 1:** List of viticultural regions and their projected risk

```{r retrieve risk table USA, echo = FALSE, message = FALSE, warning = FALSE, results = "hide"}

# use this if you want to run this vignette locally
#USA_viticultural_regions <- usa_slf_risk_report[["viticultural_regions_list"]]

# I generated the figure this way so as to show the table in the vignette
USA_viticultural_regions <- read_csv(file = file.path(here::here(), "vignette-outputs", "reports", "United States", "united_states_L_delicatula_report_viticultural_regions_list.csv")) %>%
  dplyr::slice(208:223)

USA_viticultural_regions

```


You may begin to notice that a particular region has many records, like we can see is the case for Washington State. You could then produce a report only for that region, to get a better idea of the overall trend of risk shift due to climate change. We will produce a report for Washington State alone, to better visualize this trend:

```{r generate report for washington USA, message = FALSE, warning = FALSE, results = "hide"}

scari::create_risk_report(
  locality.iso = "usa", # the country iso is still required
  locality.name = "washington", # the name must be specified
  locality.type = "state_province", # we have changed the report type
  buffer.dist = 20000, # in meters
  create.dir = FALSE,
  save.report = FALSE
)

```

We can see that most of Washington is at some level of SLF risk presently. Under climate change, risk is projected to decrease some, but one or both modeled scales still predict that SLF can establish in most of the state.

```{r retrieve current risk map for washington USA, message = FALSE, warning = FALSE, echo = FALSE}

washington_current <- washington_slf_risk_report[["risk_maps"]][["present_risk_map"]] +
  labs(
    title = element_blank(),
    subtitle = element_blank(),
    caption = element_blank()
  ) +
  theme(legend.position = "none")

washington_current

```

**Fig. 3:** Projected current  risk of *Lycorma delicatula* establishment | Washington, USA

```{r retrieve projected future risk map for washington USA, message = FALSE, warning = FALSE, echo = FALSE}

washington_future <- washington_slf_risk_report[["risk_maps"]][["future_risk_map"]] +
  labs(
    title = element_blank(),
    subtitle = element_blank(),
    caption = element_blank()
  ) +
  theme(legend.position = "none")

washington_future

```

**Fig. 4:** Projected future risk of *Lycorma delicatula* establishment under climate change | Washington, USA

Based on the viticultural risk quadrant plot, we can now see that Washington state exhibits a totally different trend from the rest of the country. While viticultural regions across the united states exhibit a range of risk levels, regions in Washington are either at high or extreme risk for SLF establishment, and this pattern does not change under predicted climate change levels.

*Note* At this scale, we can see the buffer zones used to assess viticultural region risk.

```{r retrieve risk plot washington USA, echo = FALSE, message = FALSE, warning = FALSE}

washington_risk_plot <- washington_slf_risk_report[["viticultural_risk_plot"]] +
  labs(
    title = element_blank(),
    subtitle = element_blank(),
    caption = element_blank()
  ) +
  theme(legend.position = "none")

washington_risk_plot

```

**Fig. 5:** Projected shift in the risk for *Lycorma delicatula* establishment at key viticultural regions due to climate change | Washington, USA

In our more localized analysis, we might also be interested in quantifying the total area at risk for SLF or the total number of viticultural regions at risk for SLF establishment. For this, we will look at two additional outputs from `create_risk_report()`:

1. risk map area table- this quantifies the areas and proportions of the total occupied by each risk category on the risk map
2. viticultural risk table- this quantifies the number of winegrowing regions that are projected to fall into each risk category, both now and in the future under climate change

First, let's look at the risk map area table:

**Table 2:** Table quantifying the projected current and future suitable area for *Lycorma delicatula* establishment | Washington, USA

```{r retrieve prop area table for washington USA, echo = FALSE, message = FALSE, warning = FALSE}

washington_prop_area <- washington_slf_risk_report[["risk_maps_prop_area_table"]]

washington_prop_area

```

We can see from this table that the total unsuitable area decreased under climate change, from 8.6% to 7.4%, but this change was mostly found in the regional-scale model ensemble, which predicted ~9% more suitable arae under climate change. Our modeled scales also diverged in the agreed suitable area (suitable_agreement decreased by ~9%).

Next, lets look at the viticultural risk table:

**Table 3:** Table quantifying the projected shift in the risk for *Lycorma delicatula* establishment at key viticultural regions due to climate change | Washington, USA

```{r retrieve risk table for washington USA, echo = FALSE, message = FALSE, warning = FALSE}

washington_risk_table <- washington_slf_risk_report[["viticultural_risk_table"]]

washington_risk_table

```

From this table, we can see that all 7 viticultural regions maintain their present risk level under climate change. Six regions fall into the high risk category (regional ensemble suitability only) and one falls into the extreme risk category (suitable area agreement).

Users should explore the other localities and data types available with this function.

## 2. Recreate the analysis for another invasive species of interest

### 2.1 Gather Import Data

For modelers who wish to apply this pipeline for other invasive species, this pipeline can easily be adapted to model the risk of establishment by simply changing the input datasets and modeled scales.

First, a modeler would need to change the input datasets, outlined in vignettes 020-040. In vignette 020, I retrieved input data from GBIF, which hosts datasets for thousands of other species. I recommend including extra data from other databases and the literature as I did in my analysis, but GBIF is a great starting point for data retrieval. Here is an example using the package `rgbif`. Before I perform this operation, I will need to enter my user credentials for gbif; to keep these private, I will edit the .Renviron and call them from there. Once this chunk is run and the .Renviron pops up, enter your username, password and email credentials in the following format on lines 1-3:

-   GBIF_USER=" "
-   GBIF_PWD=" "
-   GBIF_EMAIL=" "

```{r add user credentials, message = FALSE, warning = FALSE}

# edit R environment for user credentials
usethis::edit_r_environ()

```

Save and close the document, and restart R. Now I will search for species occurrence records using GBIF.

```{r retrieve SLF records using rgbif}

slf_id <- rgbif::occ_search(scientificName = "Lycorma delicatula")[["data"]]

slf_id <- slf_id %>%
  dplyr::select(taxonKey) %>%
  dplyr::slice_head() %>%
  as.character()


# initiate download
records_gbif <- rgbif::occ_download(
  # general formatting
  type = "and",
  format = "SIMPLE_CSV",
  # inclusion rules
  pred("taxonKey", slf_id), # search by GBIF ID, not species name
  pred("hasCoordinate", TRUE),
  pred("hasGeospatialIssue", FALSE),
  pred("occurrenceStatus", "PRESENT")
)

```

The input covariates would also need to change based on what is biologically relevant for the particular species. I outline the process of choosing input covariates in vignette 030.

Once the input data have been adapted to the user's needs, the regional-scale models will need to be applied per region of interest. Each regional-scale model depends on a specific background area selection for MaxEnt that will need to change. I outline the process for choosing this area in vignette 060, in which I subset the presence data by region and intersect these subsets with the Köeppen-Geiger climate zones to select the appropriate background area. This resulting polygon would need to be cropped to the region of interest. I use the `kgc`, `terra`, and `sf` packages for this analysis:

```{r import SLF presence data, echo = FALSE}

# note these data have been tidied between the download and here- see vig 020 for details
presence_data <- read_rds(file = file.path(here::here(), "data", "slf_all_coords_final_2026-06-23.rds")) %>%
  dplyr::select(-species) %>%
  sf::st_as_sf(coords = c("x", "y"), crs = "ESRI:54017")

```

```{r retrieve K-G climate zones, eval = FALSE}

# get K-G zones
kmz_data <- kgc::kmz

# generate coordinates
kmz_lat <- kgc::genCoords(latlon = "lat", full = TRUE)
kmz_lon <- kgc::genCoords(latlon = "lon", full = TRUE)

# join data and coordinates
kmz_data <- cbind(kmz_lat, kmz_lon) %>%
  cbind(., kmz_data) %>%
  as.data.frame() %>%
  # relocate column
  dplyr::select(kmz_lon, everything())

# convert to raster
KG_zones_rast <- terra::rast(
  x = kmz_data,
  type = "xyz",
  crs = "EPSG:4326"
  )

# reproject to ESRI:54017
kmz_data_rast <- terra::project(
  x = kmz_data_rast,
  y = "ESRI:54017",
  method = "bilinear", # bilinear interpolation for continuous data
  threads = TRUE # use multiple threads for faster processing
)

# next, convert raster to polygon
KG_zones_poly <- terra::as.polygons(
  x = KG_zones_rast,
  aggregate = TRUE, # combine cells with the same value into one area
  values = TRUE, # include cell values as attributes
  crs = "ESRI:54017"
)
KG_zones_poly <- sf::st_as_sf(KG_zones_poly)

# intersect polygon and presences
regional_poly <- sf::st_filter(x = KG_zones_poly, y = presence_data)

```

### 2.2 Create MaxEnt Models using specialized internal functions.

Once the presence data, covariate data and background points have been chosen, the user might apply this framework for any invasive species to create MaxEnt models for any region of interest. Here is an example using some of our built-in functions. We utilize the [SDMtune package framework](https://consbiol-unibern.github.io/SDMtune/) to prepare data for the models and train the initial models (see articles [Prepare data for the analysis](https://consbiol-unibern.github.io/SDMtune/articles/prepare-data.html) and [Train a model](https://consbiol-unibern.github.io/SDMtune/articles/train-model.html)). 

Once the models have been created, we use our functions `compute_MaxEnt_summary_statistics()`, `create_MaxEnt_suitability_maps()`, and `predict_xy_suitability()` to create various outputs. Thes functions are essentially wrappers around a group of functions from the SDMtune package, but that modifies these outputs and organizes them into an output folder. There are versions of each of these functions for Cross-validated models specifically, which end in `_CV()` (for example, `compute_MaxEnt_summary_statistics_CV()`).

Here I will show some example code for usage of our internal datasets to produce various outputs of a MaxEnt model.

```{r library packages needed for SDMtune, eval = FALSE, message = FALSE}

# SDMtune and dependencies
library(SDMtune) # main package used to run SDMs
library(dismo) # package underneath SDMtune
library(rJava) # for running MaxEnt
library(plotROC) # plots ROCs
library(viridis)

```

This first function, [compute_MaxEnt_summary_statistics()](R/compute_MaxEnt_summary_statistics.R), calculates summary statistics for the model and outputs them to a folder. This function calculates the following outputs:
* AUC and TSS goodness of fit metrics
* the ROC curve
* a confusion matrix of threshold values
* univariate and marginal response curves for each covariate
* variable permutation importance for all covariates
* jackknife leave-one-out and isolation tests for variable importance
* all generally important summary statistics for a MaxEnt Model (gain, iterations, etc)

This example was for our model of the area of east Asia invaded by SLF:
```{r compute summary statistics using custom function, eval = FALSE}

# output directory
mypath <- file.path(here::here() %>% 
                       dirname(),
                     "maxent/models/slf_regional_invaded_asian_v4")

scari::compute_MaxEnt_summary_statistics(
  model.obj = regional_invaded_asian_model, # the model object, made by SDMtune
  model.name = "regional_invaded_asian", # gives the outputs a name
  mypath = mypath, # path to the directory where it will save
  create.dir = TRUE, # create directory if needed
  env.covar.obj = x_invaded_asian_env_covariates, # env covariate rasters used to train the model
  train.obj = regional_invaded_asian_train, # training data used to create model
  test.obj = regional_invaded_asian_test, # data you wish to use to test the model
  plot.type = c("cloglog", "logistic"), # types of univariate and marginal response curves to be created
  jk.test.type = c("train", "test") # types of jackknife curves to be created
  )

```

This next function, [create_MaxEnt_suitability_maps()](R/create_MaxEnt_suitability_maps.R), will create suitability maps of the area of interest. It takes a set of rasters as an area to create an output for. It can create a raw output threshold that output by many of the common MaxEnt thresholds (MTP, MTSS, 10th percentile), and can even layer thresholded maps.

```{r predict mean and thresholded suitability maps for historical data, eval = FALSE}

scari::create_MaxEnt_suitability_maps(
  model.obj = regional_invaded_asian_model,
  model.name = "regional_invaded_asian", 
  mypath = mypath, 
  create.dir = FALSE, 
  env.covar.obj = x_global_hist_env_covariates, 
  describe.proj = "globe_1981-2010", # name of area or time period describe.proj to
  clamp.pred = TRUE, # whether or not to clamp the predictions to the range of the training data, generally a good idea
  map.thresh = TRUE, # whether thresholded versions of these maps should be created
  thresh = c("fixed_1", "10_percentile", "MTSS"), # the thresholds used to create thresholded maps, as desired
  map.thresh.extra = "fixed_1", # put an extra threshold under the main threshold in thresh
  summary.file = file.path(mypath, "regional_invaded_asian_summary.csv") # confusion matrix created by the previous function to be used in thresh
)

```

Finally, we can use the function [predict_xy_suitability()](R/predict_xy_suitability.R) to predict the suitability for a set of points based on the model. It can also output a table of the predicted suitability values for each point. It will output a .csv file of the xy coordinates with the suitability values attached. It can calculate these suitability values using either a simple or buffer method. See [terra::buffer()](https://rspatial.github.io/terra/reference/buffer.html?q=buffer#null) and [SDMtune::predict()](https://rdrr.io/r/stats/predict.html) for more details.

This chunk uses the buffer method to create predictions for a set of wineries, based on this model:
```{r predictions for IVR coordinate data- historical, eval = FALSE}

scari::predict_xy_suitability(
  xy.obj = IVR_regions,
  xy.type = "IVR regions", # descriptive name
  env.covar.obj = x_global_hist_env_covariates,
  model.obj = regional_invaded_asian_model, # the model object, made by SDMtune
  mypath = mypath,
  predict.type = c("cloglog", "logistic"), 
  clamp.pred = TRUE,
  output.name = "regional_invaded_asian_wineries_1981-2010",
  buffer.pred = TRUE,
  buffer.fun = "mean", # the function used to calculate the suitability value for the buffer area
  buffer.dist = 20000, # distance in meters
  crs = "ESRI:54017" # should be the CRS of the raster fed into the model initially
)

```

# References

Bryant, C., Wheeler, N. R., Rubel, F., & French, R. H. (2017). kgc: Koeppen-Geiger Climatic Zones. https://CRAN.R-project.org/package=kgc

Chamberlain S, Barve V, Mcglinn D, Oldoni D, Desmet P, Geffert L, Ram K (2024). rgbif: Interface to the Global Biodiversity Information Facility API. R package version 3.8.0, https://CRAN.R-project.org/package=rgbif.

Gallien, L., Douzet, R., Pratte, S., Zimmermann, N. E., & Thuiller, W. (2012). Invasive species distribution models – how violating the equilibrium assumption can create new insights. Global Ecology and Biogeography, 21(11), 1126–1136. https://doi.org/10.1111/j.1466-8238.2012.00768.x

Huron, N. A., Behm, J. E., & Helmus, M. R. (2022). Paninvasion severity assessment of a U.S. grape pest to disrupt the global wine market. Communications Biology, 5(1), 655. https://doi.org/10.1038/s42003-022-03580-w

Pebesma, E., 2018. Simple Features for R: Standardized Support for Spatial Vector Data. The R Journal 10 (1), 439-446, https://doi.org/10.32614/RJ-2018-009

Phillips, S. J., Anderson, R. P., & Schapire, R. E. (2006). Maximum entropy modeling of species geographic distributions. Ecological Modelling, 190(3), 231–259. https://doi.org/10.1016/j.ecolmodel.2005.03.026

Hijmans R (2024). terra: Spatial Data Analysis. R package version 1.7-81, https://rspatial.github.io/terra/, https://rspatial.org/.

Ushey K, Wickham H (2024). renv: Project Environments. R package version 1.0.7, https://github.com/rstudio/renv, https://rstudio.github.io/renv/.

Wickham H, Averick M, Bryan J, Chang W, McGowan LD, François R, Grolemund G, Hayes A, Henry L, Hester J, Kuhn M, Pedersen TL, Miller E, Bache SM, Müller K, Ooms J, Robinson D, Seidel DP, Spinu V, Takahashi K, Vaughan D, Wilke C, Woo K, Yutani H (2019). “Welcome to the tidyverse.” Journal of Open Source Software, 4(43), 1686. doi:10.21105/joss.01686.
