---
title: "Analyses- R package scari"
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r setup, include = FALSE}

library(scari)

```

Here is the index of articles which outlines the steps in our analysis. Start at "Initialize scari" and work through the subsequent steps to follow the package workflow.

## Step 1: [Initialize scari and dependencies](010_initialize_pkg.html)

## Step 2: Retrieve and tidy input data for MaxEnt

1. [Retrieve climatic rasters](020_retrieve_bioclim_variables.html)
2. [Retrieve occurrence data](030_retrieve_occurrence_records.html)

## Step 3: Execute SDM modeling pipeline

1. [Setup and run global-scale MaxEnt model](040_setup_and_run_global_MaxEnt_model.html)
2. [Run global-scale MaxEnt model](050_run_global_MaxEnt_model.html)
3. [Setup regional-scale MaxEnt models](060_setup_regional_MaxEnt_models.html)
4. [Run regional-scale MaxEnt model- invaded range in North America](070_run_regional_invaded_MaxEnt_model.html)
5. [Run regional-scale MaxEnt model- invaded range in Asia](080_run_regional_invaded_asian_MaxEnt_model.html)
6. [Run regional-scale MaxEnt model- native range](090_run_regional_native_MaxEnt_model.html)

## Step 4: Ensemble regional-scale SDMs

1. [Analyze extrapolation using ExDet and MIC analyses](100_run_ExDet_MIC.html)
2. [Ensemble regional-scale models using measures of extrapolation and best fit](110_ensemble_regional_models.html)

## Step 5: Quantify SLF risk

1. [Create suitability maps per model scale](120_create_suitability_maps.html)
2. [Create suitability xy plots for viticultural areas](130_create_suitability_xy_plots_viticultural.html)
3. [Create suitability xy plots for SLF occurrences](131_create_suitability_xy_plots_SLF.html)
4. [Create viticultural risk tables](160_generate_format_figures.html)

## Step 6: Validate model fit

1. [Assess biological realism of variable response curves](140_assess_response_curves.html)
2. [Assess goodness of fit via AUC, TSS and confusion matrices](141_assess_AUC_confusion_matrix.html)
3. [Assess variable importance](142_assess_var_imp.html)

### Function Workflows

- Function: [compute_MaxEnt_summary_statistics()](051_compute_MaxEnt_summary_statistics_workflow.html)
