---
title: "Create risk reports for Lycorma delicatula"
---

```{r, include = FALSE}

knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

```

```{r setup, include = FALSE}

library(scari)

```

Use these articles to create reports quantifying the risk of invasion by Lycorma delicatula (the spotted lanternfly, SLF) to viticultural regions. We provide example usage of our main function for this task, `create_risk_report()`, applied to different geographical regions.

## Quantify risk to specific viticultural regions

This section explains how to quantify the risk to specific viticultural regions using `scari`.

1. [Get started using scari](scari.html)
2. [Example usage of create_risk_report(), applied to create reports for French Viticultural areas](150_create_risk_report_example_usage.html)
3. [Create risk reports for global countries and provinces](151_create_risk_report_global.html)
4. [Create risk reports for the USA nationally and per state](152_create_risk_report_USA.html)
